### 2013-04-24 doc/icml2013/evaluation.tex
Cleaned proof of Lemma 3, and fixed the figure placements
Added comments from Percy

We implemented the low rank recovery optimization routines for $M_2$ and
$M_3$ using a proximal subgradient descent
algorithm.%\cite{candes11phaselift,tomioka2010estimation}. (Note sure about this citation and not required?)
The algorithm
was initialized with the regularized least squares solution and was run
for 500 iterations or until convergence up to a tolerance of $10^{-3}$.
%Both regularization parameters, $\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as $\frac{1}{n}$.

As a baseline, we also implemented EM with two different
initializations. In the first, we used $\beta$s with random standard
Gaussian entries and uniform mixture probabilities, $\pi$. In the other,
we initialized EM with the starting points given by Spectral Experts.

@@

@@

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=0.50\textwidth]{figures/hist.pdf}
  \caption{Histogram over recovery errors between the three algorithms. $b = 1, d = 4, k = 3, n = 500,000$}
  \label{fig:hist}
\end{figure}

>>>>

We solved the low rank recovery optimization routines for $M_2$ and
$M_3$ using \texttt{cvx}\citationneeded. 
%ainterior point method an interior proximal subgradient descent
%algorithm.%\cite{candes11phaselift,tomioka2010estimation}. (Note sure about this citation and not required?)
%The algorithm was initialized with the regularized least squares
%solution and was run for 500 iterations or until convergence up to
%a tolerance of $10^{-3}$.
%Both regularization parameters, $\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as $\frac{1}{n}$. 
As a baseline, we also implemented EM with two different
initializations. In the first, we used $\beta$s with random standard
Gaussian entries and uniform mixture probabilities, $\pi$. In the other,
we initialized EM with the starting points given by Spectral Experts.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.50\textwidth]{figures/hist.pdf}
  \caption{Histogram over recovery errors between the three algorithms. $b = 1, d = 4, k = 3, n = 500,000$}
  \label{fig:hist}
\end{figure}

@@

### 2013-04-24 doc/icml2013/intro.tex
Added comments from Percy

% Goal of this paper
% This paper: first step in establishing provably correct estimation
% with simple mixture of linear regression
Our broad goal is to develop efficient provably consistent estimators for
discriminative latent-variable models.
In this paper, we provide a first step in this 
direction by proposing a new algorithm for a simple model,
\emph{mixture of linear regressions} (MLG) \cite{VieleTong2002}.

@@

% Twist
In mixture of linear regressions, using third-order moments does not
directly reveal the tensor structure of the problem, so we cannot
directly apply the above tensor factorization techniques.  Our solution
is to employ low-rank linear regression
\cite{NegahbanWainwright2009,Tomioka2011} on second and third powers of
the response to recover the appropriate symmetric tensors, using which we
can then apply the tensor power method to retrieve the final parameters.

@@

%We have already defined $y$ to be the response variable and $x \in
%\Re^d$ to be the covariates it depends on. Note that the matrix of
%regressors $B$ is a $d \times K$ matrix, where the $k$-th column is
%$\beta_k$.

@@

We use $x\tp{p}$ to represent the $p$-th order tensor formed by taking
the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 \ldots i_p}
= x_{i_1} \cdots x_{i_p}$. We will use $\innerp{\cdot}{\cdot}$ to denote
the generalized dot product between two $p$-th order tensors:
$\innerp{X}{Y} = \sum_{i_1, \ldots i_p} X_{i_1, \ldots i_p} Y_{i_1,
\ldots i_p}$.  A tensor $X$ is symmetric if for all $i,j \in [d]^p$
which are permutations of each other, $X_{i_1 \cdots i_p}$ = $X_{j_1
\cdots j_p}$ (all tensors in this paper will be symmetric).  For
a $p$-th order tensor $X \in (\Re^d)\tp{p}$, the mode-$i$ unfolding of
$X$ is a matrix whose $j$-th row contains all the elements of $X$ whose
$i$-th index is equal to $j$. Let $X_{(i)} \in \Re^{d \times d^{p-1}}$
be the mode-$i$ unfolding of $X$.
%For $X \in (\Re^d)\tp{p}$ and $Y \in (\Re^d)\tp{(p+q)}$,% $p, q > 0$,
%$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
%$p$ modes of $Y$, i.e. $\innerp{X}{Y}_{i_1, \ldots, i_q} = \sum_{j_1,
%\ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1, \ldots,
%i_q}$.

% Norms
For a matrix $X$,
let $\|X\|_*$ denote the nuclear (trace) norm (sum of singular values),
let $\|X\|_F$ denote the Frobenius norm (square root of sum of squares of singular values),
let $\|X\|_{\max}$ denote the max norm (elementwise maximum),
let $\|X\|_\text{op}$ denote the operator norm (largest singular value).
For a tensor $X$,
let $\|X\|_* = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_*$ denote
the average nuclear norm over all $p$ unfoldings,
and let $\|X\|_\text{op} = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\text{op}$
denote the average operator norm over all $p$ unfoldings.
Let $\sigma_\text{min}(X)$ be the smallest singular value of $X$.

For a symmetric tensor $X \in (\Re^d)\tp{p}$, let $\cvec(X) \in
\Re^{\binom{d + p + 1}{p}}$ be the collapsed vectorization of distinct
elements in $X$: each component of $\cvec(X)$ is indexed by a vector of
counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.  The value
of that component is $\sum_{k \in K(c)} X_{k_1 \cdots k_p}$, where $K(c)
= \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}|
\}$ are the set of index vectors $k$ with that count profile.  For
example, for $X \in \Re^{d \times d}$, $\cvec(X) = (X_{ii} : i \in [d];
X_{ij} + X_{ji} : i,j \in [d], i<j)$.
%For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.

>>>>

% Goal of this paper
% This paper: first step in establishing provably correct estimation
% with simple mixture of linear regression
Our broad goal is to develop efficient provably consistent estimators for
discriminative latent-variable models.
In this paper, we provide a first step in this 
direction by proposing a new algorithm for a simple model,
\emph{a mixture of linear regressions} (MLG) \cite{VieleTong2002}.

@@

% Twist
In a mixture of linear regressions, using third-order moments does not
directly reveal the tensor structure of the problem, so we cannot
directly apply the above tensor factorization techniques.  Our solution
is to employ low-rank linear regression
\cite{NegahbanWainwright2009,Tomioka2011} on second and third powers of
the response to recover the appropriate symmetric tensors, using which we
can then apply the tensor power method to retrieve the final parameters.

@@

%We have already defined $y$ to be the response variable and $x \in
%\Re^d$ to be the covariates it depends on. Note that the matrix of
%regressors $B$ is a $d \times k$ matrix, where the $k$-th column is
%$\beta_k$.

@@

We use $x\tp{p}$ to represent the $p$-th order tensor formed by taking
the outer product of $x \in \Re^d$; i.e. $x\tp{p}_{i_1 \ldots i_p}
= x_{i_1} \cdots x_{i_p}$. We will use $\innerp{\cdot}{\cdot}$ to denote
the generalized dot product between two $p$-th order tensors:
$\innerp{X}{Y} = \sum_{i_1, \ldots i_p} X_{i_1, \ldots i_p} Y_{i_1,
\ldots i_p}$.  A tensor $X$ is symmetric if for all $i,j \in [d]^p$
which are permutations of each other, $X_{i_1 \cdots i_p}$ = $X_{j_1
\cdots j_p}$ (all tensors in this paper will be symmetric).  For
a $p$-th order tensor $X \in (\Re^d)\tp{p}$, the mode-$i$ unfolding of
$X$ is a matrix, $X_{(i)} \in \Re^{d \times d^{p-1}}$, whose $j$-th row contains all the elements of $X$ whose
$i$-th index is equal to $j$. 
%For $X \in (\Re^d)\tp{p}$ and $Y \in (\Re^d)\tp{(p+q)}$,% $p, q > 0$,
%$\innerp{X}{Y} \in \Re^{d^q}$ is the projection of $X$ onto the first
%$p$ modes of $Y$, i.e. $\innerp{X}{Y}_{i_1, \ldots, i_q} = \sum_{j_1,
%\ldots, j_p} X_{j_1, \ldots, j_p} Y_{j_1, \ldots, j_p, i_1, \ldots,
%i_q}$.

% Norms
For a matrix $X$,
let $\|X\|_*$ denote the nuclear (trace) norm (sum of singular values),
let $\|X\|_F$ denote the Frobenius norm (square root of sum of squares of singular values),
let $\|X\|_{\max}$ denote the max norm (elementwise maximum),
let $\|X\|_\Lop$ denote the operator norm (largest singular value),
let $\sigma_\text{min}(X)$ be the smallest singular value of $X$.
For a tensor $X$,
let $\|X\|_* = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_*$ denote
the average nuclear norm over all $p$ unfoldings,
and let $\|X\|_\Lop = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\Lop$
denote the average operator norm over all $p$ unfoldings.

For a symmetric tensor $X \in (\Re^d)\tp{p}$, let $\cvec(X) \in
\Re^{D_p}, D_p = \binom{d + p + 1}{p}$ be the collapsed vectorization of distinct
elements in $X$, for example, for $X \in \Re^{d \times d}$, $\cvec(X)
= (X_{ii} : i \in [d]; X_{ij} + X_{ji} : i,j \in [d], i<j)$.
In general, each component of $\cvec(X)$ is indexed by a vector of
counts $(c_1, \dots, c_d)$ with total sum $\sum_i c_i = p$.  The value
of that component is $\sum_{k \in K(c)} X_{k_1 \cdots k_p}$, where $K(c)
= \{ k \in [d]^p : \forall i \in [d], c_i = |\{ j \in [p] : k_j = i \}|
\}$ are the set of index vectors $k$ with that count profile.  
%For symmetric tensors, $\innerp{X}{Y} = \innerp{\cvec(X)}{\cvec(Y)}$.

### 2013-04-24 doc/icml2013/theory.tex
Added comments from Percy

\begin{theorem}[Convergence of Spectral Experts]
\label{thm:convergence}
Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d.\ points drawn from a mixture
of linear regressions model with parameter $\theta^*$
(having three independent copies simplifies the analysis).
Let $\|x\|_2 \le R$ with probability $1$
and let $\|\beta_h^*\|_2 \le L$ for all $h \in [k]$.
Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$, 
and assume $\Sigma_p \succ 0$ for each $p \in \{1,2,3\}$.
Assume $d \ge k$.
Suppose the number of samples is
$n = \max(n_1,n_2)$
where $n_1 = \Omega\left(d^6 R^{12} (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-2}) \log(d) \log(1/\delta)\right)$ and
$n_2 = \Omega\left( \sigma^6 L^6 R^{12} k (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-1}) \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
If the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
set to $\Omega\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}}\right)$,
then the parameter estimates $\hat\theta = (\hat\pi, \hat B)$ returned by
\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
satisfies 
$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\sqrt{\pi_h^*}}$.
\end{theorem}

The proof of the theorem has two parts.
First, we bound the error in the compound parameters estimates $\hat M_2,\hat M_3$
using results from \citet{Tomioka2011}.
Then, we use results from \cite{AnandkumarGeHsu2012} to convert this error
into a bound on the actual parameter estimates $\hat\theta = (\hat\pi, \hat B)$
derived from the robust tensor power method.
But first, let us study a more basic property: identifiability.

@@

We now show that some amount of unidentifiability is intrinsic to
estimation from low-order moments, not just an artifact of our
estimation procedure.  Suppose $x = (t, \dots, t^d)$.  Even if we
observed all moments $\E[x\tp{p}y\tp{p}]$ and $\E[x\tp{2p}]$ for $p \in
[r]$, all the resulting coordinates would be monomials up to degree
$2dr$, and thus the moments live in a $2(dr)$-dimensional subspace.  On
the other hand, the parameters $\theta$ live in a subspace of at least
dimension $dk$.  Therefore, at least $r \ge k/2$ moments are required
for identifiability of any algorithm for this monomial example.

@@

Our analysis is based on the low-rank regression framework of
\citet{Tomioka2011} for tensors, which builds off of
\citet{NegahbanWainwright2009} for matrices.
The main challenge is to control the noise $\eta_p(x)$,
which in addition to containing the usual Gaussian part, is also dependent on
$x$, involves the mixing noise and various cross terms.
We first set up some notation that unifies all three regressions (\refeqn{estimateM1}, \refeqn{estimateM2}, and \refeqn{estimateM3}).
Define the observation operator $\opX_p(M_p) : \Re^{d\tp{p}} \to \Re^{n}$
mapping compound parameters $M_p$:
\begin{align}
\opX_p(M_p; \sD)_i &\eqdef \innerp{M_p}{x\tp{p}_i}, & i \in [n].
\end{align}

@@

%\todo{Describe/define the convex tensor stuff.}
%We would like to exploit
%the property that $M_p$ is low rank (as typically $K \ll D$). It has
%been shown that a convex relaxation for this problem regression with
%trace norm regularization, which can be solved using a proximal
%subgradient descent algorithm\citationneeded.
%The analogue of trace norm
%regularization for higher order tensors corresponds to the sum of the
%trace norms of the mode-k unfolding of the tensor \cite{Tomioka2011},
%$X_{(k)}$, is a $d_k \times (\prod_{k' \neq k} d_{k'})$ matrix obtained
%by concatenating the entries for all dimensions other than $k$. For
%example, the 1-mode unfolding of a 3rd order tensor has entries,
%$X_{(1)}^{i_1, (i_2, i_3)} = X_{i_1, i_2, i_3}$.

%In general, the optimization problem we'd like to solve is,
%\begin{align}
%  \hat M_p &= 
%  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX_p(M_p) \|_2^2 + \frac{\lambda_n}{K} \sum_{h=1}^K \| (M_p)_{(h)} \|_* \label{eq:regression}.
%\end{align}

\begin{lemma}[\citet{Tomioka2011}, Theorem 1]
\label{lem:lowRank}
Suppose there exists a restricted strong convexity constant $\kappa(\opX_p)$ such that
$$\frac{1}{2n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX_p) \|\Delta\|^2_F \quad \text{and} \quad
\lambda_n \ge \frac{\|\opX_p^*(\eta_p)\|_{op}}{n}.$$
Then the error of $\hat M_p$ is bounded as follows:
$\| \hat M_p - M_p^* \|_F \le \frac{\lambda_n \sqrt{k}}{\kappa(\opX_p)}$.
\end{lemma}

Going forward, we need to lower bound the restricted strong convexity
parameter $\kappa(\opX_p)$ and upper bound the adjoint operator
$\|\opX_p^*(\eta_p)\|_{op}$.
%We
%will appeal to the random design framework that models the input $x$ as
%random and show bounds that hold with high probability.

%\paragraph{Adjoint operator}

%In this section, we upper bound the operator norm of the adjoint
%$\|\opX_p(\eta_p)\|_\text{op}$.

First, let us lower bound the restricted strong convexity parameter $\kappa(\opX_p)$:

\begin{lemma}[lower bound on restricted strong convexity]
\label{lem:lowRankLower}
Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
If $$n = \Omega \left( p ~ d^{2p} \log(d) \frac{R^{4p} \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
with probability at least $1-\delta$,
$$\kappa(\opX_p) \ge \frac{\sigmamin(\Sigma_p)}{4}.$$
\end{lemma}
\begin{proof}
Expanding the definition of the observation operator:
$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
= \frac{1}{2n} \sum_{(x,y) \in \sD} \innerp{\Delta}{x\tp{p}}^2$.
Unfolding the tensor and letting $\hat\Sigma_p \eqdef \frac{1}{n} \sum_{(x,y) \in \sD} \cvec(x\tp{p})\tp{2}$,
we have 
$\frac{1}{2n} \|\opX_p(\Delta)\|_2^2
= \frac{1}{2} \trace(\cvec(\Delta)\tp{2} \hat\Sigma_p)
\ge \frac{1}{2} \|\Delta\|_F^2 \sigmamin(\hat\Sigma_p)$.
It remains to show that $\sigmamin(\hat\Sigma_p)$ and $\sigmamin(\Sigma_p)$ are close.
To do this, we first apply Hoeffding's inequality elementwise.
Since $\|x\|_2 \le R$, we have that for each element $(i,j)$,
$|(\hat\Sigma_p)_{ij} - (\Sigma_p)_{ij}| = O(R^{2p}\sqrt{\frac{\log (1/\delta)}{n}})$.
Applying the union bound over the $d^{2p}$ elements of $\hat\Sigma_p - \Sigma_p$,
we have that the max norm is bounded:
$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm max} = O(R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
The max norm times $d^p$ upper bounds the Frobenius norm, which upper bounds the operator norm, so we have that
$\|\hat\Sigma_p - \Sigma_p\|_\text{\rm op} = O(d^p R^{2p} \sqrt{\frac{p \log(d) \log (1/\delta)}{n}})$.
Using the fact that $\sigmamin(\hat\Sigma_p) \ge \sigmamin(\Sigma_p) - \|\hat\Sigma_p - \Sigma_p\|_\text{\rm op}$
yields the result.
\end{proof}

Next, let us upper bound the operator norm of the observation adjoint:

\begin{lemma}[upper bound on adjoint noise]
\label{lem:lowRankUpper}
Let $\opX_p$ be the linear operator previously defined. Then with
probability at least $1-\delta$,
$$\frac1{n} \|\opX_p^*(\eta_p)\|_\op \le O\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log^3(1/\delta)}{n}}\right).$$
for each $p \in \{1,2,3\}$.
\end{lemma}
\begin{proof}
Let $\hat\E_p[f(x,\epsilon,h)]$ denote the empirical expectation over
the examples in dataset $\sD_p$ (recall the $\sD_p$'s are independent to
simplify the analysis).  By definition,
$\frac1n \|\opX_p^*(\eta_p)\|_\op = \hat\E_p[\eta_p(x) x\tp{p}]$
for $p \in \{1,2,3\}$ defined in \refeqn{y1}, \refeqn{y2}, and
\refeqn{y3}. Each $\eta_p(x)$ is composed of several zero-mean random
variables, and since $\|A + B\|_\op \le \|A\|_\op + \|B\|_\op$, it
suffices to consider each term in turn. For each term, we will
condition on $x$ and use a standard concentration inequality to bound
the noise.

%\begin{align}
%  \| \hat\E_1[ \eta_1(x) x ] \|_{op} &\le 
%  \underbrace{ \|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op }_{ O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}}) } +
%  \underbrace{ \|\hat\E_1[\epsilon x]\|_\op }_{ O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}}) } +
%\end{align}

First, consider $\eta_1(x)$, which contains two terms.
The first term is $\innerp{M_1 - \beta_h}{x}$.
We have that $\|M_1 - \beta_h\|_2 \le 2L$,
so by Hoeffding's inequality, and using the fact that $\|x\|_2 \le R$,
with probability $1-\delta_1$,
$\|\hat\E_1[\innerp{M_1 - \beta_h}{x} x]\|_\op = O(R^2 L \sqrt{\frac{\log(1/\delta_1)}{n}})$.
The second term is $\epsilon \sim \normal{0}{\sigma^2}$,
with probability $1-\delta_1$,
$\|\hat\E_1[\epsilon x]\|_\op = O(R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.

Now, consider $\eta_2(x)$, which contains three terms.
Using similar techniques, we have that
$\|\hat\E_2[\eta_2(x) x\tp{2}]\|_\op = O(R^2 (L^2 R^2 + \sigma L R + \sigma^2) \sqrt{\frac{\log(1/\delta_1)}{n}})$.

Finally, $\eta_3(x)$ contains five terms plus an additional bias since the estimator uses $\hat M_1$ rather than $M_1$.
There are two additional considerations.
First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
Second, the additional bias is small by the previous bound (and it is constructed on $\sD_1$, which is independent of $\sD_3$
used to construct $\hat M_3$.
Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
We can handle the other terms using conventional means.
Combining everything, we obtain
$\|\hat\E_3[\eta_3(x) x\tp{3}]\|_\op = O(R^3 (L^3 R^3 + \sigma L^2 R^2 + \sigma^2 L R + \sigma^3 + \sigma^3 R) \sqrt{\frac{\log^3(1/\delta_1)}{n}})$.

Finally, taking $\delta_1 = \delta/3$, and taking the union bound over the bounds for $p \in \{1,2,3\}$,
we get our result.
Note that we are bounding quantities fairly crudely for the sake of simplicity.
\end{proof}

@@

Now that we have bounds on the errors of recovering $M_2$ and $M_3$ from
the data, we will extend the analysis of the robust tensor method in
\citet{AnandkumarGeHsu2012} to bound the error on the parameters
$\theta$. 
\begin{lemma}
  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_{op} \le \epsilon$ and
  $\|\hat M_3 - M_3\|_{op} \le \epsilon$, for some $\epsilon$ such that 
  $$\epsilon < \min\{\sigmamin(M_2)/2,O(k \sigmamin(M_2)^{1.5} \|M_3\|_{op}^{-1})\}.$$ 
  Then, for some permutation of indices, the parameter estimates found
  in part 2 of \algorithmref{algo:spectral-experts} satisfy the
  following with probability atleast $1 - \delta$,
  \begin{align*}
  \|\hat \pi - \pi \|_{\infty}
    &= O( k \pi_{max}^{2.5} \pi_{min}^{-0.5} \sigmamin(M_2)^{-1.5} \|M_3\|_{op} \epsilon ) \\
  \|\hat \beta_h - \beta_h\|_2
    &= O( k \pi_{max} \sigmamin(M_2)^{-1.5} \|M_2\|_{op}^{0.5} \|M_3\|_{op} \epsilon ).
  \end{align*}
  for all $h \in [k]$.
\end{lemma}
The proof of the above lemma has been deferred to
\appendixref{sec:proofs}.

@@

\begin{proof}
By \reflem{lowRank} along with \reflem{lowRankLower} and
\reflem{lowRankUpper}, we can control the operator norm of the parameter
estimates which are directly upper bounded by the Frobenius norm: If $n
\ge n_1$, then
\begin{align}
  \|\hat M_p - M_p\|_{op} = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
  %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
\end{align}

We complete the proof by applying \reflem{tensorPower} with the above
bound on $\|\hat M_p - M_p\|_{op}$.

>>>>

\begin{theorem}[Convergence of Spectral Experts]
\label{thm:convergence}
Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d.\ points drawn from a mixture
of linear regressions model with parameter $\theta^*$
(having three independent copies simplifies the analysis).
Let $\|x\|_2 \le R$ with probability $1$
and let $\|\beta_h^*\|_2 \le L$ for all $h \in [k]$.
Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$, 
and assume $\Sigma_p \succ 0$ for each $p \in \{1,2,3\}$.
Assume $d \ge k$.
Suppose the number of samples is
$n = \max(n_1,n_2)$
where $n_1 = \Omega\left(R^{12} (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-2}) \log(1/\delta)\right)$ and
$n_2 = \Omega\left( \sigma^6 L^6 R^{12} k (\max_{p \in [3]} \sigmamin(\Sigma_p)^{-1}) \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$.
If the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
set to $\Omega\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}}\right)$,
then the parameter estimates $\hat\theta = (\hat\pi, \hat B)$ returned by
\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
satisfies 
$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\sqrt{\pi_h^*}}$.
\end{theorem}

The proof of the theorem has two parts.
First, we bound the error in the compound parameters estimates $\hat M_2,\hat M_3$
using results from \citet{Tomioka2011}.
Then, we use results from \citet{AnandkumarGeHsu2012} to convert this error
into a bound on the actual parameter estimates $\hat\theta = (\hat\pi, \hat B)$
derived from the robust tensor power method.
But first, let us study a more basic property: identifiability.

@@

We now show that some amount of unidentifiability is intrinsic to
estimation from low-order moments, not just an artifact of our
estimation procedure.  Suppose $x = (t, \dots, t^d)$.  Even if we
observed all moments $\E[x\tp{p}y\tp{p}]$ and $\E[x\tp{2p}]$ for $p \in
[r]$, all the resulting coordinates would be monomials up to degree
$2dr$, and thus the moments live in a $2dr$-dimensional subspace.  On
the other hand, the parameters $\theta$ live in a subspace of at least
dimension $dk$.  Therefore, at least $r \ge k/2$ moments are required
for identifiability of any algorithm for this monomial example.

@@

Our analysis is based on the low-rank regression framework of
\citet{Tomioka2011} for tensors, which builds off of
\citet{NegahbanWainwright2009} for matrices.
The main challenge is to control the noise $\eta_p(x)$,
which in addition to containing the usual Gaussian part, is also dependent on
$x$, involves the mixing noise and various cross terms.
We first set up some notation that unifies all three regressions (\refeqn{estimateM1}, \refeqn{estimateM2}, and \refeqn{estimateM3}).
Define the observation operator $\opX_p(M_p) : \Re^{d\tp{p}} \to \Re^{n}$
mapping compound parameters $M_p$:
\begin{align}
\opX_p(M_p; \sD)_i &\eqdef \innerp{M_p}{x\tp{p}_i}, & (x_i, y_i) \in \sD.
\end{align}

@@

%\todo{Describe/define the convex tensor stuff.}
%We would like to exploit
%the property that $M_p$ is low rank (as typically $k \ll D$). It has
%been shown that a convex relaxation for this problem regression with
%trace norm regularization, which can be solved using a proximal
%subgradient descent algorithm\citationneeded.
%The analogue of trace norm
%regularization for higher order tensors corresponds to the sum of the
%trace norms of the mode-k unfolding of the tensor \cite{Tomioka2011},
%$X_{(k)}$, is a $d_k \times (\prod_{k' \neq k} d_{k'})$ matrix obtained
%by concatenating the entries for all dimensions other than $k$. For
%example, the 1-mode unfolding of a 3rd order tensor has entries,
%$X_{(1)}^{i_1, (i_2, i_3)} = X_{i_1, i_2, i_3}$.

%In general, the optimization problem we'd like to solve is,
%\begin{align}
%  \hat M_p &= 
%  \arg\min_{M_p}& \frac{1}{2N} \| \vec y - \opX_p(M_p) \|_2^2 + \frac{\lambda_n}{k} \sum_{h=1}^k \| (M_p)_{(h)} \|_* \label{eq:regression}.
%\end{align}

\begin{lemma}[\citet{Tomioka2011}, Theorem 1]
\label{lem:lowRank}
Suppose there exists a restricted strong convexity constant $\kappa(\opX_p)$ such that
$$\frac{1}{2n} \| \opX_p( \Delta )\|_2^2 \ge \kappa(\opX_p) \|\Delta\|^2_F \quad \text{and} \quad
\lambda_n \ge \frac{\|\opX_p^*(\eta_p)\|_\Lop}{n}.$$
Then the error of $\hat M_p$ is bounded as follows:
$\| \hat M_p - M_p^* \|_F \le \frac{\lambda_n \sqrt{k}}{\kappa(\opX_p)}$.
\end{lemma}

Going forward, we need to lower bound the restricted strong convexity
parameter $\kappa(\opX_p)$ and upper bound the adjoint operator
$\|\opX_p^*(\eta_p)\|_\Lop$. The proofs of the following lemmas follow
from standard concentration inequalities, and have been deferred to
\appendixref{sec:proofs}.

%We will appeal to the random design framework that models the input $x$
%as random and show bounds that hold with high probability.
%\paragraph{Adjoint operator}
%In this section, we upper bound the operator norm of the adjoint
%$\|\opX_p(\eta_p)\|_\Lop$.

% First, let us lower bound the restricted strong convexity parameter $\kappa(\opX_p)$:

\begin{lemma}[lower bound on restricted strong convexity]
\label{lem:lowRankLower}
Let $\Sigma_p \eqdef \E[\cvec(x\tp{p})\tp{2}]$.
If $$n = \Omega \left( \frac{R^{4p} \log(1/\delta)}{\sigmamin(\Sigma_p)^2} \right),$$
with probability at least $1-\delta$,
$$\kappa(\opX_p) \ge \frac{\sigmamin(\Sigma_p)}{4}.$$
\end{lemma}

% Next, let us upper bound the operator norm of the observation adjoint:

\begin{lemma}[upper bound on adjoint noise]
\label{lem:lowRankUpper}
Let $\opX_p$ be the linear operator previously defined. Then with
probability at least $1-\delta$,
$$\frac1{n} \|\opX_p^*(\eta_p)\|_\op \le O\left(\sigma^3 L^3 R^6 \sqrt{\frac{\log^3(1/\delta)}{n}}\right).$$
for each $p \in \{1,2,3\}$.
\end{lemma}

@@

Now that we have bounds on the errors of recovering $M_2$ and $M_3$ from
the data, we will extend the analysis of the robust tensor method in
\citet{AnandkumarGeHsu2012} to include the whitening and unwhitening
steps, 2(a) and 2(c) respectively, of
\algorithmref{algo:spectral-experts}.
\begin{lemma}
  \label{lem:tensorPower} Let $\|\hat M_2 - M_2\|_\Lop \le \epsilon$ and
  $\|\hat M_3 - M_3\|_\Lop \le \epsilon$, for some $\epsilon$ such that 
  $$\epsilon < \min\{\sigmamin(M_2)/2,O(k \sigmamin(M_2)^{1.5} \|M_3\|_\Lop^{-1})\}.$$ 
  Then, there exists a permutation of indices such that  the parameter
  estimates found in step 2 of \algorithmref{algo:spectral-experts}
  satisfy the following with probability at least $1 - \delta$,
  \begin{align*}
  \|\hat \pi - \pi \|_{\infty}
    &= O( k \pi_{max}^{2.5} \pi_{min}^{-0.5} \sigmamin(M_2)^{-1.5} \|M_3\|_\Lop \epsilon ) \\
  \|\hat \beta_h - \beta_h\|_2
    &= O( k \pi_{max} \sigmamin(M_2)^{-1.5} \|M_2\|_\Lop^{0.5} \|M_3\|_\Lop \epsilon ).
  \end{align*}
  for all $h \in [k]$.
\end{lemma}

The proof follows by applying standard matrix perturbation results for
the whitening and unwhitening operators, and has again been deferred
\appendixref{sec:proofs}.

@@

\begin{proof}(Theorem 1)
By \reflem{lowRank} along with \reflem{lowRankLower} and
\reflem{lowRankUpper}, we can control the operator norm of the parameter
estimates which are directly upper bounded by the Frobenius norm: If $n
\ge n_1$, then
\begin{align}
  \|\hat M_p - M_p\|_\Lop = O\left( \sigma^3 L^3 R^6 k^{\frac12} \sigmamin(\Sigma_p)^{-1} \sqrt{\frac{\log^3 (1/\delta)}{n}} \right).
  %\|\hat M_p - M_p\|_F = O\left( \frac{\sigma^3 L^3 R^6 \sqrt{k} \log^3 (1/\delta)}{\sqrt{n} (\sigmamin(\Sigma_p) - d^p R^p \sqrt{\frac{p \log(d) \log(1/\delta)}{n}}} \right).
\end{align}

We complete the proof by applying \reflem{tensorPower} with the above
bound on $\|\hat M_p - M_p\|_\Lop$.
